---
title: "Glue numbers and characters in R"
author: "Sina Rüeger"
date: "July 13, 2020"
output:
    tint::tintHtml:
      highlight: "tango"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
library(dplyr)
library(glue)
```

## Motivation

Sometimes you want to do the following:

### 1. Write analysis results as inline text

You have a dataset, such as the following data about individuals and their current living region in Finland:

```{r echo=FALSE}
n <- 100
dat_study <- tibble(id = as.character(glue("id_{id}", id = sample(100:999, n))), state = sample(c("Uusimaa", "Ostrobothnia", "South Karelia", "Lapland"), n, replace = TRUE), height= runif(n, 150, 200))
dat_study
```

And you would like to make a statement in an R markdown file as follows: 

```{r echo = FALSE, comment = ""}
cat(htmltools::includeText("helper.Rmd"))
```

so it produces:
```{r, include = FALSE}
states <- "Ostrobothnia, Uusimaa, South Karelia and Lapland"
```

> The study participants were from four different Finnish states: `r states`. 

### 2. Custom output file path

During the course of a data analysis, we want to store out data or data visualizations using the same basename `/data/haemoglobin/experiment-2020-07-13`:

```{r, eval = FALSE}
## save a plot
ggsave("/data/plots/experiment-2020-07-13.pdf")

## save data
write_tsv(data, path = "/data/tables/experiment-2020-07-13.tsv")
```

Instead of repetition, it would be handy if we could do the following:

```{r, eval = FALSE}
dir_p <- "/data/plots/"
dir_t <- "/data/tables/"
basename <- "experiment-2020-07-13"

## save a plot
ggsave("{dir_p}{basename}.pdf")

## save data
write_tsv(data, path = "{dir_t}{basename}.tsv")
```

## Solution: Combining R data objects into strings

For these three examples, there is a solution that uses the same framework. 

`This is some text {NUMBER} and {NUMBER}.`

For example, if we define `NUMBER_SUN` and `NUMBER_SNOW`, then
```
This year, South of Finland had on average 
{NUMBER_SUN} days of sunshine and {NUMBER_SNOW} of days snowfall.
```

results in
```
This year, South of Finland had on average 
50 days of sunshine and 3 of days snowfall.
```

## Glue!

The R package [glue](https://github.com/tidyverse/glue) does exactly that. 

Here is the example from before: 
```{r, comment = "", tidy = FALSE}
glue("This year, South of Finland had on average {NUMBER_SUN} 
     days of sunshine and {NUMBER_SNOW} of days snowfall.", 
     NUMBER_SUN = 50, NUMBER_SNOW = 3)
```


If you have used the function `paste` before, then think of glue as the tidyverse answer to paste. 

## Live coding

```{r}
## install.packages("glue")
library(glue)
library(dplyr)
```

Explore all functions:
```{r}
apropos("glue")
```

We are going to look at
- `glue`
- `glue_collapse`
- `glue_data`

### Glue two elements

Glue one element (`animals`).

```{r}
animals <- "Elephants"
glue("There were {animals}.")
```

Glue two elements.

```{r}
number <- 3
glue("There were {number} {animals}.")
```

Glue two elements, but define them within the `glue`` command.
```{r}
glue("There were {number} {animals}.", number = 2, animals = "Giraffe")
```

### Exercise

Tweak the following R command so that it returns: `Yesterdays average temperature was 20.2 celsius.`

```{r, eval = FALSE}
metric <- "celsius"
temperature <- 20.2
glue("Yesterdays average temperature was [metric] [temperature].")
```

### Collapse vector

Collapse the elements in a vector

```{r}
genes <- c("FTO", "PSCK3", "DTN3")
glue_collapse(genes)
```

Collapse the elements and separate them with a `+`. 
```{r}
glue_collapse(genes, sep = " + ")
```

### Exercise

Predict the output of the following R code and choose solution 1, 2, 3 or 4 below.

```{r}
trees <- c("pine", "fir", "cedar")
glue_collapse(trees, sep = ",", last = "and")
```

1. `pine, fir and cedar`
1. `pine,firandcedar`
1. `pine,fir,cedar`
1. `pine, firandcedar`


## Fix intial problems

### 1. Write analysis results as inline text

```{r echo = FALSE, comment = ""}
cat(htmltools::includeText("helper_solution.Rmd"))
```

### 2. Combine directory and file name

Lastly, to answer problem 3) from the beginning: 


```{r, eval = FALSE}
dir_p <- "/data/plots/"
dir_t <- "/data/tables/"
basename <- "experiment-2020-07-13"

## save a plot
ggsave(glue("{dir_p}{basename}.pdf"))

## save data
write_tsv(data, path = glue("{dir_t}{basename}.tsv"))
```



## Find out more

- https://glue.tidyverse.org/index.html
- glue also appears in the [`stringr` cheatsheet](https://raw.githubusercontent.com/rstudio/cheatsheets/master/strings.pdf)
